11th International Satisfiability Modulo Theories Competition (SMT-COMP 2016): Rules and Procedures

### 2016-04-20 rules/rules16.tex
Draft of 2016 rules added.

%\usepackage{draftwatermark}

@@

\title{10th International Satisfiability Modulo Theories Competition
  (SMT-COMP 2015): Rules and Procedures}

@@

\author{%
Sylvain Conchon \\
Paris-Sud University \\
France \\
{\small\href{mailto:Sylvain.Conchon@lri.fr}{\tt Sylvain.Conchon@lri.fr}} \\
\and
David D{\'e}harbe \\
Federal University of Rio Grande do Norte \\
Brazil \\
{\small\href{mailto:david@dimap.ufrn.br}{\tt david@dimap.ufrn.br}} \\
\and
Tjark Weber \\
Uppsala University \\
Sweden \\
{\small\href{mailto:tjark.weber@it.uu.se}{\tt tjark.weber@it.uu.se}} \\
\\
}

@@

\begin{itemize}
\item SMT-COMP mailing list: \href{mailto:smt-comp@cs.nyu.edu}{\tt
  smt-comp@cs.nyu.edu}
\item Sign-up site for the mailing list:
  \url{cs.nyu.edu/mailman/listinfo/smt-comp}
\end{itemize}

\noindent Additional material will be made available at the
competition web site, \url{www.smtcomp.org} or
\url{smtcomp.sourceforge.net/2015}.

@@

\begin{description}
\item[May~1] Deadline for new benchmark contributions.
\item[June~1] Deadline for first versions of solvers (for all tracks),
  including information about which tracks and divisions are being
  entered, and magic numbers for benchmark scrambling.
\item[June~7] Final versions of competition tools (e.g., benchmark
  scrambler) are made available.  Benchmark libraries are frozen.
\item[June~14] Deadline for final versions of solvers, including
  system descriptions.
\item[June~15] Opening value of NYSE Composite Index used to complete
  random seed for benchmark scrambling.
\item[July~18/19] SMT Workshop; end of competition, presentation of
  results.
\end{description}

@@

The annual Satisfiability Modulo Theories Competition~(SMT-COMP) is
held to spur advances in SMT solver implementations on benchmark
formulas of practical interest.  Public competitions are a well-known
means of stimulating advancement in software tools.  For example, in
automated reasoning, the CASC and SAT competitions for first-order and
propositional reasoning tools, respectively, have spurred significant
innovation in their fields~\cite{leberre+03,PSS02}.  More information
on the history and motivation for SMT-COMP can be found at the
competition web site, \url{www.smtcomp.org}, and in reports on
previous
competitions~(\cite{SMTCOMP-JAR,SMTCOMP-FMSD,BDOS08,SMTCOMP-2008,SMTCOMP-2014,SMTCOMP-2012,SMTEVAL-2013}).

SMT-COMP~2015 is affiliated with the SMT Workshop~2015
(\url{http://smt2015.csl.sri.com/}), which is associated with CAV~2015
(\url{http://i-cav.org/2015/}).  The SMT Workshop and the main CAV
conference will include a block of time to present the competitors and
results of the competition.

@@

SMT-COMP 2015 will have two tracks: the conventional main track and an
application (i.e., incremental) track.  Within each track there are
multiple divisions, where each division uses benchmarks from a
specific SMT-LIB logic (or group of logics).  We will recognize
winners as measured by number of benchmarks solved; we will also
recognize solvers based on additional criteria.

The rest of this document, revised from the previous
version,\footnote{Earlier versions of this document include
  contributions from Clark Barrett, Roberto Bruttomesso, David Cok,
  David D{\'e}harbe, Morgan Deters, Alberto Griggio, Albert Oliveras,
  Aaron Stump, and Tjark Weber.} describes the rules and competition
procedures for SMT-COMP~2015.  The principal changes from the previous
competition rules are the following:
\begin{itemize}
\item All solvers will be run with $n=4$ cores available.  We will
  recognize both best sequential and best parallel performance in all
  main track divisions, using different CPU time limits.
\item The division scoring will take time spent on unsolved benchmarks
  into account.  Previous competitions only considered time spent on
  solved benchmarks.  However, the number of solved benchmarks still
  takes precedence over run-time.
\item In addition to recognizing the best solver in each division, we
  will recognize solvers that perform best according to
  competition-wide criteria, emphasizing breadth of supported logics.
  These criteria will be a refinement of the criteria used for the
  FLoC Olympic Games medals in~2014.
\item We plan to include new divisions for floating-point arithmetic.
  These should be considered experimental in~2015, and will not be
  included in the competition-wide ranking.
\item We plan to run solvers on all eligible benchmarks.  Previous
  competitions used a proper subset of eligible benchmarks, causing
  competition results to be affected by a random selection process.
\end{itemize}

@@

For participation in SMT-COMP, a solver must be uploaded to StarExec,
made publicly available, and the organizers informed of its presence
{\em and the tracks and divisions in which it is
  participating}. StarExec supports solver configurations; for
clarity, each submitted solver must have one configuration only.  A
submission must also include a 32-bit unsigned integer.  These integer
numbers, collected from all submissions, are used to seed the
benchmark scrambler.

@@

SMT-COMP entrants must be submitted via StarExec (solvers) \emph{and}
email to the organizers (accompanying information) until the end of
{\bf June~1, 2015} anywhere on earth.  After this date no new entrants
will be accepted.  However, updates to existing entrants on StarExec
will be accepted until the end of {\bf June~14, 2015} anywhere on
earth.

@@

\header{Dates of competition.}
%
The bulk of the computation will take place during the weeks leading
up to SMT~2015, from about June~15 to July~17.  Intermediate results
will be regularly posted to the SMT-COMP website as the competition
runs.

@@

\header{Input and output.}
%
In the main track, a participating solver must read a single benchmark
script, whose name is presented as the solver's first argument.  In
the application track, a trace executor will send commands from a
benchmark script to the solver's standard input channel.

The benchmark script is in the concrete syntax of the SMT-LIB format,
version~2.0, though with a restricted set of commands.  A benchmark
script is a text file containing a sequence of SMT-LIB~v2 commands in
which:
%
\begin{enumerate}
\item In the main track, there may be a (single) \akey{set-option
  :print-success ...} command.  Note that {\tt success} outputs are
  ignored by the post-processor used by the
  competition.\footnote{SMT-LIB~2.0 requires that a solver produce a
    \texttt{success} answer after each \akey{set-logic},
    \akey{declare-sort}, \akey{declare-fun} and \akey{assert} command
    (among others), unless the option \akey{:print-success} is set to
    false.  Ignoring the \texttt{success} outputs allows for
    submitting fully SMT-LIB~2.0 compliant solvers without the need
    for a wrapper script, while still allowing entrants of previous
    competitions to run without changes.}
\item In the application track, the \akey{:print-success} option may
  not be disabled; the application track executor will send an initial
  $\akey{set-option :print-success true}$ command to the solver.
\item The (single) $\akey{set-logic}$ command setting the benchmark's
  logic is the first command after any $\akey{set-option}$ commands.
\item The script next optionally contains (any number of)
  $\akey{set-info}$ commands, which may be ignored by the
  solver. Benchmarks will likely have some of the following
  \akey{set-info} commands: \akey{(set-info :smt-lib-version 2.0)},
  \akey{(set-info :category ...)}, \akey{(set-info :source ...)}, and
  \akey{(set-info :status ...)}.
\item In the main track, there is exactly one $\akey{check-sat}$
  command, following possibly several $\akey{assert}$ commands.
\item In the application track, there are one or more \akey{check-sat}
  commands, each preceded by one or more \akey{assert} commands.
  There may also be zero or more \akey{push 1} commands, and zero or
  more \akey{pop 1} commands, consistent with the uses of those
  commands in the SMT-LIB standard. Each \akey{check-sat} command may
  be preceded by a \akey{set-info :status} command that indicates the
  expected result of the \akey{check-sat} command.
\item The script may optionally contain an $\akey{exit}$ command as
  its last command.  In the application track, this command must not
  be omitted.
\item The formulas in the script belong to the benchmark's logic, with
  any free symbols declared in the script.
\item All sorts declared with a $\akey{declare-sort}$ command must
  have zero arity.
\item Application track scripts may have $\akey{define-sort}$
  commands.
\item No other commands besides the ones just mentioned may be used.
\item Named formulas are not used in benchmark scripts.
\end{enumerate}
%
The SMT-LIB format specification is available from the ``Documents''
section of the SMT-LIB website~\cite{SMT-LIB}.  Solvers will be given
formulas only from the divisions into which they have been entered.

@@

The main track competition will consist of selected benchmarks in each
of the logic divisions.  Each benchmark script will be presented to
the solver as its first command-line argument.  Each SMT-COMP entrant
is then expected to attempt to report on its standard output channel
whether the formula is satisfiable (\texttt{sat}, in lowercase) or
unsatisfiable (\texttt{unsat}).  An entrant may also report
\texttt{unknown} to indicate that it cannot determine satisfiability
of the formula.

The main track competition uses a StarExec post-processor (named
``SMT-COMP 2015") to accumulate the results.

\header{Aborts and unparsable output.}
%
Any \texttt{success} outputs will be ignored.  Solvers that exit
before the time limit without reporting a result (e.g., due to
exhausting memory or crashing) and do not produce output that includes
\texttt{sat}, \texttt{unsat} or \texttt{unknown} will be considered to
have aborted.

\header{Persistent state.}
%
Solvers may create and write to files and directories during the
course of an execution, but they must not read such files back during
later executions.  Each solver is executed with a temporary directory
as its current working directory.  Any generated files should be
produced there (and not, say, in the system's {\tt /tmp} directory).
The StarExec system sets a limit on the amount of disk storage
permitted---typically 20\,GB.  See the StarExec user guide for more
information.  The temporary directory is deleted after the job is
complete.  Solvers must not attempt to communicate with other
machines, e.g., over the network.

@@

The application track evaluates SMT solvers when interacting with an
external verification framework, e.g., a model checker. This
interaction, ideally, happens by means of an online communication
between the model checker and the solver: the model checker repeatedly
sends queries to the SMT solver, which in turn answers either
\texttt{sat} or \texttt{unsat}.  In this interaction an SMT-solver is
required to accept queries incrementally via its {\em standard input
  channel}.

In order to facilitate the evaluation of the solvers in this track, we
will set up a ``simulation'' of the aforementioned interaction. In
particular each benchmark in the application track represents a
realistic communication trace, containing multiple \akey{check-sat}
commands (possibly with corresponding \akey{push 1}/\akey{pop 1}
commands), which is parsed by a \emph{trace executor}. The trace
executor serves the following purposes:
\begin{itemize}
\item it simulates the online interaction by sending single queries to
  the SMT solver (through stdin);
\item it prevents ``look-ahead'' behaviors of SMT solvers;
\item it records time and answers for each call;
\item it guarantees a fair execution for all solvers by abstracting
  from any possible crash, misbehavior, etc. that may happen on the
  model checker side.
\end{itemize}
%
The trace executor terminates processing the benchmark script upon
receiving an incorrect response from the solver.

@@

\header{Input and output.}
%
Participating solvers will be connected to a trace executor, which
will incrementally send commands to the standard input channel of the
solver and read responses from the standard output channel of the
solver.  The commands will be taken from an SMT-LIB~2.0 benchmark
script that satisfies the rules for an application script given in
Section~\ref{sec:logistics}.

Solvers must respond to each command sent by the trace executor with
the answers defined in the SMT-LIB~2.0 format specification, that is,
with a \texttt{success} answer for \akey{set-option},
\akey{set-logic}, \akey{declare-sort}, \akey{declare-fun},
\akey{define-sort}, \akey{define-fun}, \akey{assert}, \akey{push 1},
and \akey{pop 1} commands, and with an answer of \texttt{sat},
\texttt{unsat}, or \texttt{unknown} for \akey{check-sat} commands.
There will be no \akey{get-unsat-core}, \akey{get-proof},
\akey{get-model}, \akey{get-assignments}, or \akey{get-values}
commands in application track benchmarks.

@@

\header{Benchmark sources.}
%
Benchmarks for each division will be drawn from the SMT-LIB benchmark
repository.  The main track will use a subset of all
\emph{non-incremental} benchmarks; the application track will use a
subset of all \emph{incremental} benchmarks.

\header{New benchmarks.}
%
The deadline for submission of new benchmarks is {\bf May~1, 2015}.
The organizers will be checking and curating these until {\bf June~7,
  2015}.  The organizers reserve the right to exclude new benchmarks
if any prove problematic for some reason.  SMT-COMP attempts to give
preference to benchmarks that are ``real-world,'' in the sense of
coming from or having some intended application outside SMT.

@@

\header{Benchmark selection.}
%
The benchmark pool is culled as follows:
\begin{enumerate}
\item \emph{Retire inappropriate benchmarks.} The competition
  organizers may remove from the eligibility pool certain benchmarks
  that are inappropriate\footnote{In 2015, this (again) includes all
    benchmarks that use bit-vector division or remainder, because of
    an unresolved discussion about the semantics of division by zero.}
  or uninteresting for competition, or cut the size of certain
  benchmark families to avoid their over-representation.
\item \emph{Eliminate benchmarks whose status is unknown.} Any
  benchmark whose expected output is neither {\tt sat} nor {\tt unsat}
  is removed.  There are a significant number of benchmarks with {\tt
    unknown} status in the library; these are certainly interesting
  and are a challenge to solve, but they are not used in the
  competition.\footnote{Incremental benchmarks may contain multiple
    \akey{check-sat} commands, each with its own status.  Incremental
    benchmarks that contain (some) \akey{check-sat} commands with {\tt
      unknown} status may still contain a prefix of commands where the
    status is known for all \akey{check-sat} commands in the prefix.
    This prefix is eligible for the application track.}
\end{enumerate}
%
All remaining benchmarks are used for the competition.  There will be
no further selection of benchmarks, e.g., based on benchmark
difficulty or benchmark category.

@@

\section{Scoring}

@@

Divisions for floating-point arithmetic are experimental in 2015, and
will not be considered competitive.

@@

\header{Main track.} More specifically, for the main track, we have
%
\begin{itemize}
\item $e=0$, $n=0$ if the solver times out, or the solver aborts
  without a response, or the result of the \akey{check-sat} command is
  {\tt unknown},
\item $e=0$, $n=1$ if the solver does not time out, and the result of
  the \akey{check-sat} command is correct,
\item $e=1$, $n=0$ if the solver does not time out, and the result of
  the \akey{check-sat} command is incorrect.
\end{itemize}
%
Note that for a (correct or incorrect) result to be taken into
consideration, the solver process must terminate (normally or
abnormally) within the time limit.

@@

SMT-COMP has traditionally emphasized sequential performance (i.e.,
CPU time) over parallel performance (i.e., wall-clock time).  StarExec
is measuring both, and we intend to recognize both best sequential and
best parallel solvers in all competitive main track divisions.

@@

To compute a solver's score for a division, the solver's individual
benchmark scores for all benchmarks in the division are summed
component-wise.  For main track divisions, we will separately compute
the sum of all raw scores (Section~\ref{sec:benchmark-scoring}) to
assess parallel performance, and the sum of all sequential scores
(Section~\ref{sec:sequential}) to assess sequential performance.  For
application track divisions, division scores will be based on raw
scores.\footnote{Since application track benchmarks may be partially
  solved, defining a useful sequential score for the application track
  would require information not provided by the raw score, e.g.,
  detailed timing information for each result.}

Division scores are compared lexicographically:
\begin{itemize}
\item A sum of raw scores $\langle e, n, w, c\rangle$ is better than
  $\langle e', n', w', c'\rangle$ iff $e < e'$ or ($e = e'$ and $n >
  n'$) or ($e = e'$ and $n = n'$ and $w < w'$) or ($e = e'$ and $n =
  n'$ and $w = w'$ and $c < c'$).  That is, fewer errors takes
  precedence over more correct solutions, which takes precedence over
  less wall-clock time taken, which takes precedence over less CPU
  time taken.
\item A sum of sequential scores $\langle e_S, n_S, c_S\rangle$ is
  better than $\langle e_S', n_S', c_S'\rangle$ iff $e_S < e_S'$ or
  ($e_S = e_S'$ and $n_S > n_S'$) or ($e_S = e_S'$ and $n_S = n_S'$
  and $c_S < c_S'$).  That is, fewer errors takes precedence over more
  correct solutions, which takes precedence over less CPU time taken.
\end{itemize}
%
We will not make any comparisons between raw scores and sequential
scores, as these are intended to measure fundamentally different
performance characteristics.

\subsection{Competition-wide scoring}

@@

The organizers will also recognize the following contributions:
%
\begin{itemize}
\item \emph{Open source.}  The top solver that provides its source
  code as open source will be recognized in each competitive division.
\item \emph{Best industrial performance.}  There is agreement in the
  SMT community to emphasize problems that come from real
  applications.  The solver that performs best on industrial
  benchmarks, as measured by the division scoring restricted to this
  benchmark category, will be recognized in each competitive division.
\item {\em Best new entrant}. The best performing entrant from a new
  solver implementation team, as measured by the competition-wide
  metric.
\item {\em Breadth of logics}. Solvers that cover the most theories
  and logics.
\item {\em Benchmarks}. Contributors of new benchmarks.
\end{itemize}
%
The organizers reserve the right to recognize other outstanding
contributions that become apparent in the competition results.

@@

>>>>

\usepackage{draftwatermark}

@@

\title{11th International Satisfiability Modulo Theories Competition
  (SMT-COMP 2016): Rules and Procedures}

@@

\author{%
Sylvain Conchon \\
Paris-Sud University \\
France \\
{\small\href{mailto:Sylvain.Conchon@lri.fr}{\texttt{Sylvain.Conchon@lri.fr}}} \\
\and
David D{\'e}harbe \\
Clearsy, France \& Federal University\\ of Rio Grande do Norte, Brazil \\
{\small\href{mailto:david.deharbe@clearsy.com}{\texttt{david.deharbe@clearsy.com}}} \\
\and
Tjark Weber \\
Uppsala University \\
Sweden \\
{\small\href{mailto:tjark.weber@it.uu.se}{\texttt{tjark.weber@it.uu.se}}} \\
\\
}

@@

\begin{itemize}
\item SMT-COMP mailing list:
  \href{mailto:smt-comp@cs.nyu.edu}{\texttt{smt-comp@cs.nyu.edu}}
\item Sign-up site for the mailing list:
  \url{cs.nyu.edu/mailman/listinfo/smt-comp}
\end{itemize}

\noindent Additional material will be made available at the
competition web site, \url{www.smtcomp.org} or
\url{smtcomp.sourceforge.net/2016}.

@@

\begin{description}
\item[May~1] Deadline for new benchmark contributions.
\item[May~15] Deadline for first versions of solvers (for all tracks),
  including information about which tracks and divisions are being
  entered, and magic numbers for benchmark scrambling.
\item[May~22] Final versions of competition tools (e.g., benchmark
  scrambler) are made available.  Benchmark libraries are frozen.
\item[May~29] Deadline for final versions of solvers, including
  system descriptions.
\item[May~30] Opening value of NYSE Composite Index used to complete
  random seed for benchmark scrambling.
\item[July~1/2] SMT Workshop; end of competition, presentation of
  results.
\end{description}

@@

The annual Satisfiability Modulo Theories Competition~(SMT-COMP) is
held to spur advances in SMT solver implementations on benchmark
formulas of practical interest.  Public competitions are a well-known
means of stimulating advancement in software tools.  For example, in
automated reasoning, the CASC and SAT competitions for first-order and
propositional reasoning tools, respectively, have spurred significant
innovation in their fields~\cite{leberre+03,PSS02}.  More information
on the history and motivation for SMT-COMP can be found at the
competition web site, \url{www.smtcomp.org}, and in reports on
previous
competitions~(\cite{SMTCOMP-JAR,SMTCOMP-FMSD,BDOS08,SMTCOMP-2008,CDW14,SMTCOMP-2012,CSW15}).

SMT-COMP~2016 is affiliated with the SMT Workshop~2016
(\url{http://smt-workshop.cs.uiowa.edu/2016/}), which is associated
with IJCAR~2016 (\url{http://ijcar-2016.info/}).  The SMT Workshop
will include a block of time to present the results of the
competition.

@@

SMT-COMP 2016 will have two tracks: the conventional main track and an
application (i.e., incremental) track.  Within each track there are
multiple divisions, where each division uses benchmarks from a
specific SMT-LIB logic (or group of logics).  We will recognize
winners as measured by number of benchmarks solved; we will also
recognize solvers based on additional criteria.

The rest of this document, revised from the previous
version,\footnote{Earlier versions of this document include
  contributions from Clark Barrett, Roberto Bruttomesso, David Cok,
  Sylvain Conchon, David D{\'e}harbe, Morgan Deters, Alberto Griggio,
  Albert Oliveras, Aaron Stump, and Tjark Weber.} describes the rules
and competition procedures for SMT-COMP~2016.  The principal changes
from the previous competition rules are the following:
\begin{itemize}
\item Benchmarks will use (a subset of) version~2.5 of the SMT-LIB
  language.  \emph{Rationale:} This is the latest version of the
  SMT-LIB language.  It was released on 2015-06-28, and is largely
  backwards-compatible to version~2.0, which was used in the 2015 SMT
  Competition.
\item Solver output may affect the score even when the solver does not
  terminate within the time limit.  (In 2015, main track solver output
  was ignored if the solver subsequently timed out.)  Solvers should
  take care not to accidentally produce output that contains
  \texttt{sat} or \texttt{unsat} even when they are killed.
  \emph{Rationale:} Users are likely to trust solver responses even
  when the solver continues to run for some time.
\item Divisions for floating-point arithmetic are no longer
  experimental, and will be considered competitive if the necessary
  requirements (see Section~\ref{sec:scoring}) are met.
  \emph{Rationale:} Floating-point divisions were experimental in
  2015.  By now, their definition is sufficiently stable, and they are
  supported by several solvers.
\item Division scores will be based on a weighted sum of scores for
  benchmark families.  \emph{Rationale:} For some years now, the SMT
  Competition has had sufficient computational resources to evaluate
  all solver entrants on all eligible benchmarks.  As a side-effect,
  the weighting of benchmark families that was achieved in early
  SMT-COMPs through a careful selection of benchmarks (based on
  benchmark difficulty and category) was lost.  We believe that a
  careful weighting (see Section~\ref{sec:division-scoring}), which
  de-emphasizes large benchmark families, will lead to more meaningful
  competition results.
\item Best industrial performance will no longer be recognized
  separately.  \emph{Rationale:} While there is agreement in the SMT
  community to emphasize problems that come from real applications,
  industrial performance largely coincided with overall performance in
  previous SMT-COMPs (except in some divisions that had a very small
  number of industrial benchmarks).  Therefore, industrial performance
  has very limited significance, and we consider the overhead to
  determine and report it separately not justified.
\end{itemize}

@@

For participation in SMT-COMP, a solver must be uploaded to StarExec,
made publicly available, and the organizers informed of its presence
\emph{and the tracks and divisions in which it is participating}.
StarExec supports solver configurations; for clarity, \emph{each
  submitted solver must have one configuration only}.  A submission
must also include a 32-bit unsigned integer.  These integer numbers,
collected from all submissions, are used to seed the benchmark
scrambler.

@@

SMT-COMP entrants must be submitted via StarExec (solvers) \emph{and}
email to the organizers (accompanying information) until the end of
{\bf May~15, 2016} anywhere on earth.  After this date no new entrants
will be accepted.  However, updates to existing entrants on StarExec
will be accepted until the end of {\bf May~29, 2016} anywhere on
earth.

@@

\header{Dates of competition.}
%
The bulk of the computation will take place during the weeks leading
up to SMT~2016, from about May~30 to June~30.  Intermediate results
will be regularly posted to the SMT-COMP website as the competition
runs.

@@

\header{Input and output.}
%
In the main track, a participating solver must read a single benchmark
script, whose filename is presented as the solver's first command-line
argument.  In the application track, a trace executor will send
commands from a benchmark script to the solver's standard input
channel.

The benchmark script is in the concrete syntax of the SMT-LIB format,
version~2.5, though with a restricted set of commands.  A benchmark
script is a text file containing a sequence of SMT-LIB commands that
satisfies the following requirements:
%
\begin{enumerate}
\item In the main track, there may be a (single) \akey{set-option
  :print-success ...} command.  Note that \texttt{success} outputs are
  ignored by the post-processor used by the
  competition.\footnote{SMT-LIB~2.5 requires solvers to produce a
    \texttt{success} answer after each \akey{set-logic},
    \akey{declare-sort}, \akey{declare-fun} and \akey{assert} command
    (among others), unless the option \akey{:print-success} is set to
    false.  Ignoring the \texttt{success} outputs allows for
    submitting fully SMT-LIB~2.5 compliant solvers without the need
    for a wrapper script, while still allowing entrants of previous
    competitions to run without changes.}
\item In the application track, the \akey{:print-success} option must
  not be disabled.  The trace executor will send an initial
  $\akey{set-option :print-success true}$ command to the solver.
\item The (single) $\akey{set-logic}$ command setting the benchmark's
  logic is the first command after any $\akey{set-option}$ commands.
\item The script may contain any number of $\akey{set-info}$ commands.
  These may be ignored by solvers.
\item The script may contain any number of $\akey{declare-sort}$
  commands.  All sorts declared with a $\akey{declare-sort}$ command
  must have zero arity.
\item Application track scripts may contain $\akey{define-sort}$
  commands.  \textcolor{red}{TODO: Some main track benchmarks also
    appear to contain \akey{define-sort} and \akey{define-fun}
    commands.}
\item The script may contain any number of $\akey{declare-fun}$
  commands.  All formulas in the script belong to the benchmark's
  logic, with any free symbols declared in the script.
\item In the main track, there is exactly one $\akey{check-sat}$
  command, following possibly several $\akey{assert}$ commands.
\item In the application track, there are one or more \akey{check-sat}
  commands, each preceded by one or more \akey{assert} commands.
  There may also be zero or more \akey{push 1} commands, and zero or
  more \akey{pop 1} commands, consistent with the use of those
  commands in the SMT-LIB standard.
\item The script may optionally contain an $\akey{exit}$ command as
  its last command.  In the application track, this command must not
  be omitted.
\item No other commands besides the ones just mentioned may be used.
\item Named formulas are not used in benchmark scripts.
\end{enumerate}
%
The SMT-LIB format specification is available from the ``Standard''
section of the SMT-LIB website~\cite{SMT-LIB}.  Solvers will be given
formulas only from the divisions into which they have been entered.

@@

The main track competition will consist of selected benchmarks in each
of the logic divisions.  Each benchmark script will be presented to
the solver as its first command-line argument.  The solver is then
expected to attempt to report on its standard output channel whether
the formula is satisfiable (\texttt{sat}, in lowercase) or
unsatisfiable (\texttt{unsat}).  A solver may also report
\texttt{unknown} to indicate that it cannot determine satisfiability
of the formula.

The main track competition uses a StarExec post-processor (named
``SMT-COMP 2016") to accumulate the results.

\header{Aborts and unparsable output.}
%
Any \texttt{success} outputs will be ignored.  Solvers that exit
before the time limit without reporting a result (e.g., due to
exhausting memory or crashing) \emph{and} do not produce output that
includes \texttt{sat}, \texttt{unsat} or \texttt{unknown} will be
considered to have aborted.

\header{Persistent state.}
%
Solvers may create and write to files and directories during the
course of an execution, but they must not read such files back during
later executions.  Each solver is executed with a temporary directory
as its current working directory.  Any generated files should be
produced there (and not, say, in the system's \texttt{/tmp}
directory).  The StarExec system sets a limit on the amount of disk
storage permitted---typically 20\,GB.  See the StarExec user guide for
more information.  The temporary directory is deleted after the job is
complete.  Solvers must not attempt to communicate with other
machines, e.g., over the network.

@@

The application track evaluates SMT solvers when interacting with an
external verification framework, e.g., a model checker. This
interaction, ideally, happens by means of an online communication
between the framework and the solver: the framework repeatedly sends
queries to the SMT solver, which in turn answers either \texttt{sat}
or \texttt{unsat}.  In this interaction an SMT solver is required to
accept queries incrementally via its \emph{standard input channel}.

In order to facilitate the evaluation of solvers in this track, we
will set up a ``simulation'' of the aforementioned interaction.  Each
benchmark in the application track represents a realistic
communication trace, containing multiple \akey{check-sat} commands
(possibly with corresponding \akey{push 1} and \akey{pop 1} commands),
which is parsed by a \emph{trace executor}. The trace executor serves
the following purposes:
\begin{itemize}
\item it simulates the online interaction by sending single queries to
  the SMT solver (through stdin);
\item it prevents ``look-ahead'' behaviors of SMT solvers;
\item it records time and answers for each command;
\item it guarantees a fair execution for all solvers by abstracting
  from any possible crash, misbehavior, etc.\ that might happen in the
  verification framework.
\end{itemize}
%
The trace executor terminates processing the benchmark script upon
receiving an incorrect response from the solver.

@@

\header{Input and output.}
%
Participating solvers will be connected to a trace executor, which
will incrementally send commands to the standard input channel of the
solver and read responses from the standard output channel of the
solver.  The commands will be taken from an SMT-LIB benchmark script
that satisfies the requirements for application track scripts given in
Section~\ref{sec:logistics}.

Solvers must respond to each command sent by the trace executor with
the answers defined in the SMT-LIB format specification, that is, with
an answer of \texttt{sat}, \texttt{unsat}, or \texttt{unknown} for
\akey{check-sat} commands, and with a \texttt{success} answer for
other commands.

@@

\header{Benchmark sources.}
%
Benchmarks for each division will be drawn from the SMT-LIB benchmark
library.  The main track will use a subset of all
\emph{non-incremental} benchmarks; the application track will use a
subset of all \emph{incremental} benchmarks.

\header{New benchmarks.}
%
The deadline for submission of new benchmarks is {\bf May~1, 2016}.
The organizers will be checking and curating these until {\bf May~22,
  2016}.  The organizers reserve the right to exclude new benchmarks
if any prove problematic for some reason.  SMT-COMP attempts to give
preference to benchmarks that are ``real-world,'' in the sense of
coming from or having some intended application outside SMT.

@@

\header{Benchmark selection.}
%
The benchmark pool is culled as follows:
\begin{enumerate}
\item \emph{Retire inappropriate benchmarks.} The competition
  organizers may remove from the eligibility pool certain benchmarks
  that are inappropriate\footnote{In 2016, this (again) includes all
    benchmarks that use partial or underspecified functions, e.g.,
    \texttt{bvudiv}, \texttt{fp.min}.}  or uninteresting for
  competition, or cut the size of certain benchmark families to avoid
  their over-representation.
\item \emph{Eliminate benchmarks whose status is unknown.} Any
  benchmark whose expected output is neither \texttt{sat} nor
  \texttt{unsat} is removed.  There are a significant number of
  benchmarks with \texttt{unknown} status in the library; these are
  certainly interesting and are a challenge to solve, but they are not
  used in the competition.\footnote{Incremental benchmarks may contain
    multiple \akey{check-sat} commands, each with its own status.
    Incremental benchmarks that contain (some) \akey{check-sat}
    commands with \texttt{unknown} status may still contain a prefix
    of commands where the status is known for all \akey{check-sat}
    commands in the prefix.  This prefix is eligible for the
    application track.}
\end{enumerate}
%
All remaining benchmarks are used for the competition.  There will be
no further selection of benchmarks, e.g., based on benchmark
difficulty or benchmark category.

@@

\section{Scoring}
\label{sec:scoring}

@@

@@

\header{Main track.} More specifically, for the main track, we have
%
\begin{itemize}
\item $e=0$, $n=0$ if the solver aborts without a response, or the
  result of the \akey{check-sat} command is \texttt{unknown},
\item $e=0$, $n=1$ if the result of the \akey{check-sat} command is
  correct,
\item $e=1$, $n=0$ if the result of the \akey{check-sat} command is
  incorrect.
\end{itemize}
%
Note that a (correct or incorrect) response is taken into
consideration even when the the solver process terminates abnormally,
or does not terminate within the time limit.  Solvers should take care
not to accidentally produce output that contains \texttt{sat} or
\texttt{unsat}.

@@

SMT-COMP has traditionally emphasized sequential performance (i.e.,
CPU time) over parallel performance (i.e., wall-clock time).  StarExec
measures both, and we intend to recognize both best sequential and
best parallel solvers in all competitive main track divisions.

@@

To compute a solver's score for a division, the solver's individual
benchmark scores for all benchmarks in the division are first
multiplied by a scalar weight that depends on the benchmark's family,
and then summed component-wise.

For a given competition benchmark~$b$, let~$F_b \geq 1$ be the total
number of benchmarks in $b$'s benchmark family that were used in the
competition.  We define the weight for benchmark~$b$ as $\alpha_b = (1
+ \log F_b) / F_b$.\footnote{See Section~\ref{sec:competition-wide}
  for a motivating discussion of log scaling.}  We define the
\emph{normalized weight} for benchmark~$b$ as $\alpha'_b = \alpha_b /
(\sum_{b'} \alpha_{b'})$, where the sum is over all benchmarks in the
division.

For main track divisions, we will separately compute the weighted sum
of all raw scores (Section~\ref{sec:benchmark-scoring}) $$\sum_b
\alpha'_b \cdot \langle e_b, n_b, w_b, c_b\rangle$$ where the sum is
over all benchmarks in the division to assess parallel performance,
and the weighted sum of all sequential scores
(Section~\ref{sec:sequential}) to assess sequential performance.  For
application track divisions, division scores will be based on raw
scores only.\footnote{Since application track benchmarks may be
  partially solved, defining a useful sequential score for the
  application track would require information not provided by the raw
  score, e.g., detailed timing information for each result.}

Division scores are compared lexicographically:
\begin{itemize}
\item A weighted sum of raw scores $\langle e, n, w, c\rangle$ is
  better than $\langle e', n', w', c'\rangle$ iff $e < e'$ or ($e =
  e'$ and $n > n'$) or ($e = e'$ and $n = n'$ and $w < w'$) or ($e =
  e'$ and $n = n'$ and $w = w'$ and $c < c'$).  That is, fewer errors
  takes precedence over more correct solutions, which takes precedence
  over less wall-clock time taken, which takes precedence over less
  CPU time taken.
\item A weighted sum of sequential scores $\langle e_S, n_S,
  c_S\rangle$ is better than $\langle e_S', n_S', c_S'\rangle$ iff
  $e_S < e_S'$ or ($e_S = e_S'$ and $n_S > n_S'$) or ($e_S = e_S'$ and
  $n_S = n_S'$ and $c_S < c_S'$).  That is, fewer errors takes
  precedence over more correct solutions, which takes precedence over
  less CPU time taken.
\end{itemize}
%
We will not make any comparisons between raw scores and sequential
scores, as these are intended to measure fundamentally different
performance characteristics.

\subsection{Competition-wide scoring}
\label{sec:competition-wide}

@@

The organizers will also recognize the following contributions:
%
\begin{itemize}
\item \emph{Open source.}  The top solver that provides its source
  code as open source will be recognized in each competitive division.
\item \emph{Best new entrant}. The best performing entrant from a new
  solver implementation team, as measured by the competition-wide
  metric.
\item \emph{Breadth of logics}. Solvers that cover the most theories
  and logics.
\item \emph{Benchmarks}. Contributors of new benchmarks.
\end{itemize}
%
The organizers reserve the right to recognize other outstanding
contributions that become apparent in the competition results.

@@

%% Local Variables:
%% mode: latex
%% mode: flyspell
%% ispell-local-dictionary: "american"
%% LocalWords: Conchon logics satisfiability SMT Sylvain Tjark
%% End:
